---
title: "Playoffs 2 - SPREAD Prediction Models"
author: "Janice Kim"
date: "3/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tibble)
```

```{r}
games = read.csv("games_4_7.csv")
summary(games)
```
##### Spread = Home Points−Away Points
Use pace variable for these models

## Prep Work: Creating Spread & Exploratory Analysis
```{r, eval = FALSE}
head(games)
```
Looking at histograms and plots of bwt versus each of the  explanatory variables. 
```{r, eval = FALSE}
games = games %>%
  transform(SPREAD = PTS_home - PTS_away)

names(games)
```

```{r}
Y<- games$SPREAD
X1 <- games$FG_PCT_home
X2 <- games$FT_PCT_home
X3 <- games$FG3_PCT_home
X4 <- games$AST_home
X5 <- games$REB_home
X6 <- games$FG_PCT_away
X7 <- games$FT_PCT_away
X8 <- games$FG3_PCT_away
X9 <- games$AST_away
X10 <- games$REB_away
X11 <- games$OREB_home
X12 <- games$FGA_home
X13 <- games$FGM_home
X14 <- games$FG3A_home
X15 <- games$FG3M_home
X16 <- games$OREB_away
X17 <- games$FGA_away
X18 <- games$FGM_away
X19 <- games$FG3A_away
X20 <- games$FG3M_away
X21 <- games$WSE_home
X22 <- games$WSE_away
X23 <- games$WSE_DIFF
X24 <- games$SEASON_PACE_home
X25 <- games$SEASON_PACE_away
X26 <- games$SEASON_ORB._home
X27 <- games$SEASON_ORB._away
```

```{r, eval = FALSE}
full.lm = lm(Y~X1+X2+X3+X4+X5+X6+X7+X8+X9+X10+X11+X12+X13+X14+X15+X16+X17+X18+X19+X20+X21+X22+X23+X24+X25+X26+X27)
summary(full.lm)
```


Conducted an ANOVA test to see if at least one of the coefficients is significantly different from zero.
```{r, eval = F}
anova(full.lm)
```
```
Significantly large p-value indicates that it is not a good predictor for Y.
Can remove X25 - X27

Significantly low p-value indicates that these may be significantly different from zero.
X1 - X24
```

## Method 1: Backward Elimination
First wrote updated model without variables X25 - X27 and saw that the adjusted R2 value is the same as the full model, further proving that variables x25 - x27 are not useful in predicting spread.

```{r}
str(games$SPREAD)
```

```{r}
summary(lm(games$SPREAD~X1+X2+X3+X4+X5+X6+X7+X8+X9+X10+X11+X12+X13+X14+X15+X16+X17+X18+X19+X20+X21+X22+X23+X24)) #0.8453 
```
```
Next, set the significance at alpha = 0.05 and deleted variables with the highest p-value above a until all predictor variables hve p value under 0.05.

Variable X1 has the greatest p-value at 0.91492 so we will run a model without variable X1.
```

```{r}
summary(lm(Y~X2+X3+X4+X5+X6+X7+X8+X9+X10+X11+X12+X13+X14+X15+X16+X17+X18+X19+X20+X21+X22+X23+X24))
```

```
X3 has the greatest p-value above alpha at 0.9651 so remove X3.
```

```{r}
summary(lm(Y~X2+X4+X5+X6+X7+X8+X9+X10+X11+X12+X13+X14+X15+X16+X17+X18+X19+X20+X21+X22+X23+X24))
```
```
Remove X21 at 0.59985 
```

```{r}
summary(lm(Y~X2+X4+X5+X6+X7+X8+X9+X10+X11+X12+X13+X14+X15+X16+X17+X18+X19+X20+X22+X23+X24))
```
```
Remove X4 with p-value 0.43779
```

```{r}
summary(lm(Y~X2+X5+X6+X7+X8+X9+X10+X11+X12+X13+X14+X15+X16+X17+X18+X19+X20+X22+X23+X24))
```
```
Remove X23 with p-value of 0.61426
```

```{r}
summary(lm(Y~X2+X5+X6+X7+X8+X9+X10+X11+X12+X13+X14+X15+X16+X17+X18+X19+X20+X22+X24))
```
```
Remove X14 with p-value of 0.486484
```

```{r}
summary(lm(Y~X2+X5+X6+X7+X8+X9+X10+X11+X12+X13+X14+X15+X16+X17+X18+X19+X20+X22+X24))
```
```
Remove X14 with p-value of 0.486484 
```

```{r}
summary(lm(Y~X2+X5+X6+X7+X8+X9+X10+X11+X12+X13+X15+X16+X17+X18+X19+X20+X22+X24))
```
```
Remove X8 with a p-value of 0.441259.
```

```{r}
summary(lm(Y~X2+X5+X6+X7+X9+X10+X11+X12+X13+X15+X16+X17+X18+X19+X20+X22+X24))
```
```
Remove X22 with a p-value of 0.441259.
```

```{r}
summary(lm(Y~X2+X5+X6+X7+X9+X10+X11+X12+X13+X15+X16+X17+X18+X19+X20+X24))
```
```
Remove X9, which has a p-value of 0.168817
```

```{r}
summary(lm(Y~X2+X5+X6+X7+X10+X11+X12+X13+X15+X16+X17+X18+X19+X20+X24))
```
```
Remove X19 with a p-value of 0.061562
```

```{r}
summary(lm(Y~X2+X5+X6+X7+X10+X11+X12+X13+X15+X16+X17+X18+X20+X24))
```
```
Remove X6
```

```{r}
summary(lm(Y~X2+X5+X7+X10+X11+X12+X13+X15+X16+X17+X18+X20+X24))
```

Then we need to look at the adjusted R-squared. A higher adjusted R-squared indicates that the predictor variables add predictive value. 

```
From looking at the adjusted R-squared values of the different predictive models, it can be determined that a 13-variable model is least favorable, with an adjusted R-squared of 0.8452. 
The best model is a 16-variable model, which had an adjusted R-squared value of 0.8454.

Our 16-variable model is as follows:
```

```{r}
final.lm = lm(Y~X2+X5+X7+X10+X11+X12+X13+X15+X16+X17+X18+X20+X24)
```
```
The variables in this model to predict Y (aka Spread) are:
X2:FT_PCT_home
X5:REB_home
X7:FT_PCT_away
X10:REB_away
X11:OREB_home
X12:FGA_home
X13:FGM_home
X15:FG3M_home
X16:OREB_away
X17:FGA_away
X18:FGM_away
X20:FG3M_away
X24:SEASON_PACE_home
```


## Method 2: Using RegSubsets

```{r}
library("leaps")

model_sel <- regsubsets(SPREAD~FG_PCT_home+FT_PCT_home+FG3_PCT_home+AST_home+REB_home+FG_PCT_away+FT_PCT_away+FG3_PCT_away+AST_away+REB_away+OREB_home+FGA_home+FGM_home+FG3A_home+FG3M_home+OREB_away+FGA_away+FGM_away+FG3A_away+FG3M_away+SEASON_PACE_home, data = games, nbest = 1, nvmax = 13, method = "backward")

plot(model_sel, scale="adjr2")
```
```
FT_PCT_home
REB_home
FT_PCT_away
REB_away
OREB_home
FGA_home
FGM_home
FG3M_home
OREB_away
FGA_away
FGM_away
FG3M_away
SEASON_PACE_home
```


```{r}
plot(model_sel, scale="bic")
```
```
From the model selection based on adjusted R-squared and BIC, we can see that the best 16-variable models uses the same variables as the one we found via backwards elimination.
```

```{r}
summary(final.lm)
```
```
Note that the adjusted r-squared dropped by 0.0001. Is it worth testing for collinearity?
```


## Method 3: Cross Validation

What is the purpose of cross validation?
```
Cross validation alllows you to test your model on the sample before finalizing it.
```

```{r}
install.packages("caret")
library(caret)
```


```{r}
sample.size = floor(0.8*nrow(games))
sample.size

set.seed(101)

sample = sample.int(n =nrow(games), size = floor(0.8*nrow(games)), replace = F)
train = games[sample, ]
test = games[-sample, ]
```



```{r}
ctrl = trainControl(method = "cv", number = 10)

model = train(SPREAD~FT_PCT_home+REB_home+FT_PCT_away+REB_away+OREB_home+FGA_home+FGM_home+FG3M_home+OREB_away+FGA_away+FGM_away+FG3M_away+SEASON_PACE_home,data = games, method = "lm", trControl = ctrl)

print(model)
```

```
RMSE = 5.5185
Rsquared = 0.8453
MAE = 4.4075

The lower the RMSE and MAE and the higher Rsquared, the more closely a model can predict the actual observations.
```

```{r}
model$finalModel
```

```
SPREAD = 20.112*FCT_PCT_home - 0.072*REB_home - 19.5*80FT_PCT_away + 0.084*REB_away + 0.710*OREB_home - 0.598*FGA_home + 1.882*FGM_home + 0.844*FG3M_home - 0.726*OREB_away + 0.5875*FGA_away - 1.876*FGM_away - 0.84*FG3M_away - 0.108*SEASON_PACE_home + 10.975
```

```{r}
#testing the model with the test dataset
test
```
```{r}
#For game 3...
game3.SPREAD = (20.11*0.371) - (0.072*45) - (19.58*0.629) + (0.084*48) + (0.71*20) - (0.598*105) + (1.882*39) + (0.844*11) - (0.726*9) + (0.5875*75) - (1.876*40) - (0.84*31) - (0.108*98.9) + 10.975
game3.SPREAD
```

